---
title: "P1_Social_Engagement"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(tidyverse, lme4, ggunchained, wesanderson)

setwd("C:/Users/victm/Dropbox/CogSci 4/Comp modeling/Eye Tracking/portfolio 1/Portfolio-1-Eyetracking")

```

Welcome to SOCIAL ENGAGEMENT,
this markdown is divided into following sections:
-preprocessing
-models
-results


# PREPROCESSING
Based on code from Fabio's day 4 Rmd

Outline:
a) working on raw data
-Loading raw data from the scanner
-Clean data (standardize dominant eye)
-Plotting raw data, 2 plots
-Event estimation
-Plotting raw data, while marking estimated events, 4 plots

b) working on event data
-Create Fixations and Saccades datasets / Load in Fixations and Saccards from EyeLink

## A)
### loading data in right format
```{r}
# loading data
Samples_raw <- read_csv('data/samples2018_2019.csv') %>%
  filter(Task == "SocialEngagement")

# variable types
Samples_raw <- Samples_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Trial = as.factor(Trial),
    DominantEye = ifelse(is.na(RightEye_PositionX), 'Left', 'Right'),
    RightEye_PositionX = as.numeric(RightEye_PositionX),
    RightEye_PositionY = as.numeric(RightEye_PositionY),
    RightEye_PupilSize = as.numeric(RightEye_PupilSize),
    RightEye_MeanVelocityX = as.numeric(RightEye_MeanVelocityX),
    RightEye_MeanVelocityY = as.numeric(RightEye_MeanVelocityY),
    RightEye_MeanAccellerationX = as.numeric(RightEye_MeanAccellerationX),
    RightEye_MeanAccellerationY = as.numeric(RightEye_MeanAccellerationY),
    LeftEye_PositionX = as.numeric(LeftEye_PositionX),
    LeftEye_PositionY = as.numeric(LeftEye_PositionY),
    LeftEye_PupilSize = as.numeric(LeftEye_PupilSize),
    LeftEye_MeanVelocityX = as.numeric(LeftEye_MeanVelocityX),
    LeftEye_MeanVelocityY = as.numeric(LeftEye_MeanVelocityY),
    LeftEye_MeanAccellerationX = as.numeric(LeftEye_MeanAccellerationX),
    LeftEye_MeanAccellerationY = as.numeric(LeftEye_MeanAccellerationY),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),     
    SearchOrder = as.factor(SearchOrder),
    # sampling rate "is"" 1000 Hz, we're converting to miliseconds
    TrialTime = as.numeric(TrialTime*0.001),
    TotTrialLengthInSeconds = as.numeric(TotTrialLengthInSeconds)
  )
```


### cleaning of Samples
```{r}
Samples <- Samples_raw %>%
  # convert right- and left- data into a single column
  # note: position is in pixels, [0, 0] would be the bottom left corner
  # some machines do it differently
  mutate(
    PositionX = ifelse(DominantEye == 'Right', RightEye_PositionX, LeftEye_PositionX),
    PositionY = ifelse(DominantEye == 'Right', RightEye_PositionY, LeftEye_PositionY),
    PupilSize = ifelse(DominantEye == 'Right', RightEye_PupilSize, LeftEye_PupilSize),
    MeanVelocityX = ifelse(DominantEye == 'Right', RightEye_MeanVelocityX, LeftEye_MeanVelocityX),
    MeanVelocityY = ifelse(DominantEye == 'Right', RightEye_MeanVelocityY, LeftEye_MeanVelocityY),
    MeanAccellerationX = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationX, LeftEye_MeanAccellerationX),
    MeanAccellerationY = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationY, LeftEye_MeanAccellerationY)
  ) %>% 
  # clear out the mess
  select(-c(RightEye_PositionX, RightEye_PositionY, RightEye_PupilSize, RightEye_MeanVelocityX, RightEye_MeanVelocityY, RightEye_MeanAccellerationX, RightEye_MeanAccellerationY, LeftEye_PositionX, LeftEye_PositionY, LeftEye_PupilSize, LeftEye_MeanVelocityX, LeftEye_MeanVelocityY, LeftEye_MeanAccellerationX, LeftEye_MeanAccellerationY))

rm(Samples_raw)
```


### plotting raw data
--> what is the use/interpretation of this plot. 
```{r}
x = subset(Samples, ParticipantID == '2019-2019-5_1_f2')
ggplot(x, aes(x = PositionX, y = PositionY, color = TrialTime)) +
  geom_point(size = 0.3) +
  ggtitle(paste('Sample data, subject:', x$ParticipantID, ', task:', x$Task )) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point() +
  scale_color_gradientn(colours = wes_palette("Zissou1", 21, type = "continuous"))
```


### estimation of events
--> we need to make sure that our values are also acceptable from a litterature point of view. I.e., Fabio said that velocity cut-off for saccade had been reported as between 15-50 (right now we have 10 which might be somewhat low). 
--> generally we made the parameters less conservative based on visual inspection. 

```{r}
Samples2 <- Samples %>% 
  group_by(ParticipantID, Trial) %>%
  ## if velocity is < 20 degrees/sec, the eye is "not moving"
  mutate(
    # velocity goes from minus to plus depending on the direction, ORIGINAL = -20, 20. 
    LowVelocity = ifelse(
      as.numeric(MeanVelocityX)>-5 & #changed to 15. 
      as.numeric(MeanVelocityX)<5, 1, 0)
  ) %>%
  # assume movement during the first 9 rows (purely to make the code work)
  mutate(LowVelocity = coalesce(LowVelocity, 0)) %>%
  
  ## create rolling window for averaging, ORIGINAL = 20
  mutate(Event = (cumsum(LowVelocity) - lag(cumsum(LowVelocity), n = 4))/4) %>%
  
  ## the next lines are very hacky
  ## if average is > .9, count as fixation
  ## ORIGINAL = 0.75
  mutate(FixationNum = ifelse(Event >= 0.75, 1, 0),
         .count = ifelse(c(0, diff(FixationNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         FixationNum = ifelse(FixationNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  
  ## otherwise count as saccade
  mutate(SaccadeNum = ifelse(Event < 0.90, 1, 0),
         .count = ifelse(c(0, diff(SaccadeNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         SaccadeNum = ifelse(SaccadeNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  mutate(Event = ifelse(Event >= 0.90, 'Fixation', 'Saccade'))
```


### plotting event estimation: spatal
--> also something about what we are saying here. 
```{r}
# PLOT FOR TRIAL 1
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & Trial == 1 & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Sample data with event estimation, trial', x$Trial)) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  theme_janco_point()
  

# FACET PLOT FOR EVERY TRIAL
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Sample data with event estimation, all trials') +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


### plotting event estimation: temporal & spatial
--> write something here as well (probably). 
```{r}
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & Trial == 6 & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.6) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Time sequence plot, trial', x$Trial)) +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  theme_janco_point()

ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.5) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Time sequence plot, all trials') +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


## B
### load EyeLink data (blackbox parameters)
```{r}
# load in csvs
Fixations_raw <- read_csv('data/fixations2018_2019.csv')
Saccades_raw <- read_csv('data/saccades2018_2019.csv')

# variable types
Fixations_EL <- Fixations_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

Saccades_EL <- Saccades_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Direction = as.factor(Direction),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

rm(Fixations_raw, Saccades_raw)
```



### Fixations & Saccades from Samples2 (own parameters)
I (Victor) have added a couple to the fixation to make it smooth with the logfiles. 
```{r}

# FIXATIONS
Fixations <- Samples2 %>% 
  group_by(ParticipantID, Trial, FixationNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionX = median(PositionX),
    PositionY = median(PositionY),
    Year = Year[1],
    PupilSize = median(PupilSize) #added. 
  ) %>% 
  filter(!is.na(FixationNum))

# SACCADES
Saccades <- Samples2 %>% 
  group_by(ParticipantID, Trial, SaccadeNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    SearchOrder = SearchOrder[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionXStart = min(PositionX),
    PositionXEnd = max(PositionX),
    AmplitudeX = max(PositionX)-min(PositionX),
    MeanVelocityX = median(MeanVelocityX),
    MeanAccellerationX = median(MeanAccellerationX),
    PositionYStart = min(PositionY),
    PositionYEnd = max(PositionY),
    AmplitudeY = max(PositionY)-min(PositionY),
    MeanVelocityY = median(MeanVelocityY),
    MeanAccellerationY = median(MeanAccellerationY),
    Year = Year[1] #added year
  ) %>% 
  filter(!is.na(SaccadeNum))
```

Merging logfiles and Fixations / Saccades
Now we are looking at a subset of the data (2018).
Do we agree that Riccardo/Fabio didn't find logfiles for 2019?

```{r}

# FIXATIONS 
# Fixations "Trial" column --> numeric for compatibility. 
Fixations$Trial <- as.numeric(Fixations$Trial)

#loading logfiles
logfiles <- read.csv("data/logfilesSE.csv")

#deleting 2018
Fixations$ParticipantID <- substring(Fixations$ParticipantID, 6, 20)

# SOCIAL ENGAGMENT
Fixations_SE_raw <- logfiles %>%
  full_join(Fixations, by = c("Task", "Trial", "ParticipantID")) %>%
  filter(Year == "2018") 

Fixations_SE <- Fixations_SE_raw %>%
  mutate(stim_person = ifelse(str_detect(Video, "m_"), "Male", "Female"),
         direct = ifelse(str_detect(Video, "dir"), "Direct", "Non-direct"),
         ost = ifelse(str_detect(Video, "_\\+o"), "Ostentive", "Non-ostentive")) 

#SACCADES 
#deleting 2018
Saccades$ParticipantID <- substring(Saccades$ParticipantID, 6, 20)

#Trial to numeric
Saccades$Trial <- as.numeric(Saccades$Trial)

# SOCIAL ENGAGMENT
Saccades_SE_raw <- logfiles %>%
  full_join(Saccades, by = c("Task", "Trial", "ParticipantID")) %>%
  filter(Year == "2018")
  #deleted the 2018 thing because that is 

Saccades_SE <- Saccades_SE_raw %>%
  mutate(stim_person = ifelse(str_detect(Video, "m_"), "Male", "Female"),
         direct = ifelse(str_detect(Video, "dir"), "Direct", "Non-direct"),
         ost = ifelse(str_detect(Video, "_\\+o"), "Ostentive", "Non-ostentive")) 

```

H1 preprocessing (Fixations_SE)

```{r}
#Plot showing that something is weird. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#Which person is weird? 
which(Fixations_SE$StartTime > 7) #participantID 2018-1_2_f

#Max StartTime excluding the outlier 
Fixations_SE %>% 
  filter(!ParticipantID == "2018-1_2_f") %>%
  summarize(max(StartTime))

#Excluding every datapoint which has StartTime > 6.325
Fixations_SE <- Fixations_SE %>%
  filter(!StartTime > 6.325)

#Checking 
max(Fixations_SE$StartTime) #now excluded

#Plot showing that something is weird. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#New column (condition)
Fixations_SE <- Fixations_SE %>%
  mutate(condition = paste0(direct, ost)) 

#Recoding to numbers
Fixations_SE$condition <- as.factor(as.numeric(as.factor(Fixations_SE$condition)))

#plot
ggplot(Fixations_SE, aes(condition, PupilSize)) +
  geom_violin() +
  geom_boxplot(width = .1)
  #direct & non-ostensive = 1 
  #direct & ostensive = 2
  #non-direct & non-ostensive = 3
  #non-direct & ostensive = 4


```


# H1 MODEL (PUPIL SIZE)
This block is for the main hypothesis:
pupil size (dilation) by ostension and direction.
participants can vary on stim_person, ost * direction, StartTime & StartTime^2

```{r}

#Maximal interaction model with everything.  
Model1 <- glmer(PupilSize ~ ost * direct + StartTime + stim_person * ParticipantGender + #include this
            (1 + stim_person | ParticipantID) + #gender, familiarity individual diff
            (1 + ost * direct | ParticipantID) + #
            (1 + StartTime | ParticipantID), #center StartTime around 0 & linear 
            family = gaussian(link=log), 
            data = Fixations_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(Model1) #nothing. 

#plot of participants (randomly thrown in there)
ggplot(Fixations_SE, aes(StartTime, PupilSize, fill = ParticipantID)) +
  geom_smooth(aes(color = ParticipantGender), method = "lm") #terrible, would be nice to have comp. 

#Only modelling StartTime (just to see)
Model1_remix <- glmer(PupilSize ~ ost * direct + stim_person * ParticipantGender + StartTime +
            (1 + StartTime | ParticipantID), 
            family = gaussian(link=log), 
            data = Fixations_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(Model1_remix) #very significant stuff. 

#Only ost * direction, everything else modelled away. 
Model_remix2 <- glmer(PupilSize ~ ost * direct +
            (1 + StartTime | ParticipantID) + #does this make sense?
            (1 + stim_person | ParticipantID) + #does this make sense?
            (1 + ParticipantGender | ParticipantID), #does this make sense?
            family = gaussian(link=log), 
            data = Fixations_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(Model_remix2) #super significant main effects & significant interaction. 

```

H1 PLOTS (PUPIL SIZE)
These plots are connected with the main hypothesis: 
pupil size by ostension and directedness. 

```{r}

#differences in pupilsize by ostension & directedness (main hyp)
#shows very different baselines (but otherwise not super informative.)
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, 
                         color = ost)) +
  geom_smooth() +
  facet_grid(ParticipantID~direct) +
  ylim(800,3000) #limits that are sensible. 

#Suggesting that Ostensive is "slower". We need a growth-curve to explain. 
#Also ostensive does seem to follow a consistently different trend than non-ost. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#pupilsize by ostension & directedness (weird. that ostensive = less pupilsize.)
#Direct does seem to affect pupil size (especially in ostensive, interaction?)
dodge <- position_dodge(width = 0.5)
ggplot(Fixations_SE, aes(x = ost, y = PupilSize, fill = direct)) +
  geom_violin(position = dodge)+
  geom_boxplot(width=.1, outlier.colour = NA, position = dodge) #come back. geom_jitter


```

H2: Fixation duration 

```{r}
#duration of fixation. 

Model2 <- glmer(Duration ~ ost * direct * (StartTime + I(StartTime^2)) +
            (1 + stim_person | ParticipantID) + #gender, familiarity
            (1 + ost * direct | ParticipantID) + #
            (1 + StartTime + I(StartTime^2) | ParticipantID), 
            family = gaussian(link=log), 
            data = Fixations_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(Model2)
```

H2 Plots

```{r}

```


H3: Saccades
preprocessing (Saccades_SE)

```{r}

#Making a mean velocity column based on x & y velocity. 
Saccades_SE <- Saccades_SE %>%
  mutate(MeanVelocity = sqrt(MeanVelocityX^2 + MeanVelocityY^2))

#Same participant has longer trial (of course)
ggplot(Saccades_SE, aes(x = StartTime, y = MeanVelocity, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#Max StartTime excluding the outlier 
Saccades_SE %>% 
  filter(!ParticipantID == "2018-1_2_f") %>%
  summarize(max(StartTime))

#Excluding every datapoint which has StartTime > 6.426
Saccades_SE <- Saccades_SE %>%
  filter(!StartTime > 6.426)

#Checking 
max(Saccades_SE$StartTime) #now excluded

#Plot showing that something is weird. 
ggplot(Saccades_SE, aes(x = StartTime, y = MeanVelocity, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#New column (condition)
Saccades_SE <- Saccades_SE %>%
  mutate(condition = paste0(direct, ost)) 

#Recoding to numbers
Saccades_SE$condition <- as.factor(as.numeric(as.factor(Saccades_SE$condition)))

#plot
ggplot(Saccades_SE, aes(condition, MeanVelocity)) +
  geom_boxplot()
  #direct & non-ostensive = 1 
  #direct & ostensive = 2
  #non-direct & non-ostensive = 3
  #non-direct & ostensive = 4

```

H2 MODELS (Saccades_SE)
This block is for the other hypothesis:
Saccade Velocity by ostension and direction (& start time).
I will try both interaction between ostension & direction, & without interaction. 

```{r}
# calculate saccade velocity. 
library(tidyverse)

#Model interaction
Model3 <- glmer(MeanVelocity ~ ost * direct +
            (1 + stim_person | ParticipantID) + 
            (1 + ost * direct | ParticipantID) +
            (1 + StartTime | ParticipantID), 
            family = gaussian(link=log), 
            data = Saccades_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(Model3) #nothing significant 


```

H3: Plots for mean velocity 

```{r}
dodge <- position_dodge(width = 0.5)

#longer tails in ostensive & direct
ggplot(Saccades_SE, aes(x = ost, y = MeanVelocity, fill = direct)) +
  geom_violin(position = dodge)+
  geom_boxplot(width=.1, outlier.colour = NA, position = dodge)

#no difference in median. 
p1 <- Saccades_SE %>%
  filter(MeanVelocity < 50) %>%
  ggplot(aes(x = ost, y = MeanVelocity, fill = direct)) +
  geom_violin(position = dodge)+
  geom_boxplot(width=.1, outlier.colour = NA, position = dodge)

p1

ggplot(Saccades_SE, aes(x = StartTime, y = MeanVelocity, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

```

H4: number of saccades (Saccades_SE)

```{r}

```

H4: plots of number of saccades (Saccades_SE)

```{r}

```


Plots for bogus-hypothesis: 
1. gender
2. familiarity. 

```{r}
#pupilsize by gender of participant & gender of stim person. 
#Also, not very informative besdies differing baselines
#Perhaps there is something about male_part. having steeper slope for female_pres.
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ParticipantGender)) +
  geom_smooth() +
  facet_grid(ParticipantID~stim_person)

#pupilsize by gender averaged.
#Generally males slopes are steeper (for both female & male) but especially for female - so there might be something there. Also females have slightly steeper slope for Andreas. Looks completely fucked. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ParticipantGender)) +
  geom_smooth(method = "loess") + 
  facet_grid(~stim_person) #somewhat weird w. crazy participant to the left. 

#pupilsize by gender, hmm.. female generally higher?
#Generally larger pupilsize for female_presenter. Female_participants more spread out than males. Male_participants generally bigger pupil size. 
ggplot(Fixations_SE, aes(x = ParticipantGender, y = PupilSize, fill = stim_person))+
  geom_violin(position = dodge) +
  geom_boxplot(width = .1, outlier.colour = NA, position = dodge)

```



# RESULTS

```{r}

```

