---
title: "P1_Social_Engagement"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
devtools:install_github("jancounchained/ggunchained")
devtools:install_github("jancounchained/wesanderson")
p_load(tidyverse, lme4, lmerTest, ggunchained, wesanderson, sjPlot, sjmisc, sjlabelled, ggridges)

setwd("C:/Users/victm/Dropbox/CogSci 4/Comp modeling/Eye Tracking/portfolio 1/Portfolio-1-Eyetracking")

```

Welcome to SOCIAL ENGAGEMENT,
this markdown is divided into following sections:
-preprocessing
-models
-results


# PREPROCESSING
Based on code from Fabio's day 4 Rmd

Outline:
a) working on raw data
-Loading raw data from the scanner
-Clean data (standardize dominant eye)
-Plotting raw data, 2 plots
-Event estimation
-Plotting raw data, while marking estimated events, 4 plots

b) working on event data
-Create Fixations and Saccades datasets / Load in Fixations and Saccards from EyeLink

## A)
### loading data in right format
```{r}
# loading data
Samples_raw <- read_csv('data/samples2018_2019.csv') %>%
  filter(Task == "SocialEngagement")

# variable types
Samples_raw <- Samples_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Trial = as.factor(Trial),
    DominantEye = ifelse(is.na(RightEye_PositionX), 'Left', 'Right'),
    RightEye_PositionX = as.numeric(RightEye_PositionX),
    RightEye_PositionY = as.numeric(RightEye_PositionY),
    RightEye_PupilSize = as.numeric(RightEye_PupilSize),
    RightEye_MeanVelocityX = as.numeric(RightEye_MeanVelocityX),
    RightEye_MeanVelocityY = as.numeric(RightEye_MeanVelocityY),
    RightEye_MeanAccellerationX = as.numeric(RightEye_MeanAccellerationX),
    RightEye_MeanAccellerationY = as.numeric(RightEye_MeanAccellerationY),
    LeftEye_PositionX = as.numeric(LeftEye_PositionX),
    LeftEye_PositionY = as.numeric(LeftEye_PositionY),
    LeftEye_PupilSize = as.numeric(LeftEye_PupilSize),
    LeftEye_MeanVelocityX = as.numeric(LeftEye_MeanVelocityX),
    LeftEye_MeanVelocityY = as.numeric(LeftEye_MeanVelocityY),
    LeftEye_MeanAccellerationX = as.numeric(LeftEye_MeanAccellerationX),
    LeftEye_MeanAccellerationY = as.numeric(LeftEye_MeanAccellerationY),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),     
    SearchOrder = as.factor(SearchOrder),
    # sampling rate "is"" 1000 Hz, we're converting to miliseconds
    TrialTime = as.numeric(TrialTime*0.001),
    TotTrialLengthInSeconds = as.numeric(TotTrialLengthInSeconds)
  )
```


### cleaning of Samples
```{r}
Samples <- Samples_raw %>%
  # convert right- and left- data into a single column
  # note: position is in pixels, [0, 0] would be the bottom left corner
  # some machines do it differently
  mutate(
    PositionX = ifelse(DominantEye == 'Right', RightEye_PositionX, LeftEye_PositionX),
    PositionY = ifelse(DominantEye == 'Right', RightEye_PositionY, LeftEye_PositionY),
    PupilSize = ifelse(DominantEye == 'Right', RightEye_PupilSize, LeftEye_PupilSize),
    MeanVelocityX = ifelse(DominantEye == 'Right', RightEye_MeanVelocityX, LeftEye_MeanVelocityX),
    MeanVelocityY = ifelse(DominantEye == 'Right', RightEye_MeanVelocityY, LeftEye_MeanVelocityY),
    MeanAccellerationX = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationX, LeftEye_MeanAccellerationX),
    MeanAccellerationY = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationY, LeftEye_MeanAccellerationY)
  ) %>% 
  # clear out the mess
  select(-c(RightEye_PositionX, RightEye_PositionY, RightEye_PupilSize, RightEye_MeanVelocityX, RightEye_MeanVelocityY, RightEye_MeanAccellerationX, RightEye_MeanAccellerationY, LeftEye_PositionX, LeftEye_PositionY, LeftEye_PupilSize, LeftEye_MeanVelocityX, LeftEye_MeanVelocityY, LeftEye_MeanAccellerationX, LeftEye_MeanAccellerationY))

rm(Samples_raw)
```


### plotting raw data
--> what is the use/interpretation of this plot. 
```{r}
x = subset(Samples, ParticipantID == '2019-2019-5_1_f2')
ggplot(x, aes(x = PositionX, y = PositionY, color = TrialTime)) +
  geom_point(size = 0.3) +
  ggtitle(paste('Sample data, subject:', x$ParticipantID, ', task:', x$Task )) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point() +
  scale_color_gradientn(colours = wes_palette("Zissou1", 21, type = "continuous"))
```


### estimation of events
--> we need to make sure that our values are also acceptable from a litterature point of view. I.e., Fabio said that velocity cut-off for saccade had been reported as between 15-50 (right now we have 10 which might be somewhat low). 
--> generally we made the parameters less conservative based on visual inspection. 

```{r}
Samples2 <- Samples %>% 
  group_by(ParticipantID, Trial) %>%
  ## if velocity is < 20 degrees/sec, the eye is "not moving"
  mutate(
    # velocity goes from minus to plus depending on the direction, ORIGINAL = -20, 20. 
    LowVelocity = ifelse(
      as.numeric(MeanVelocityX)>-5 & #changed to 15. 
      as.numeric(MeanVelocityX)<5, 1, 0)
  ) %>%
  # assume movement during the first 9 rows (purely to make the code work)
  mutate(LowVelocity = coalesce(LowVelocity, 0)) %>%
  
  ## create rolling window for averaging, ORIGINAL = 20
  mutate(Event = (cumsum(LowVelocity) - lag(cumsum(LowVelocity), n = 4))/4) %>%
  
  ## the next lines are very hacky
  ## if average is > .9, count as fixation
  ## ORIGINAL = 0.75
  mutate(FixationNum = ifelse(Event >= 0.75, 1, 0),
         .count = ifelse(c(0, diff(FixationNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         FixationNum = ifelse(FixationNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  
  ## otherwise count as saccade
  mutate(SaccadeNum = ifelse(Event < 0.90, 1, 0),
         .count = ifelse(c(0, diff(SaccadeNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         SaccadeNum = ifelse(SaccadeNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  mutate(Event = ifelse(Event >= 0.90, 'Fixation', 'Saccade'))
```


### plotting event estimation: spatal
--> also something about what we are saying here. 
```{r}
# PLOT FOR TRIAL 1
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & Trial == 1 & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Sample data with event estimation, trial', x$Trial)) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  theme_janco_point()
  

# FACET PLOT FOR EVERY TRIAL
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Sample data with event estimation, all trials') +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


### plotting event estimation: temporal & spatial
--> write something here as well (probably). 
```{r}
ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & Trial == 6 & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.6) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Time sequence plot, trial', x$Trial)) +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  theme_janco_point()

ggplot(subset(Samples2, ParticipantID == '2019-2019-5_1_f2' & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.5) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Time sequence plot, all trials') +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


## B
### load EyeLink data (blackbox parameters)
```{r}
# load in csvs
Fixations_raw <- read_csv('data/fixations2018_2019.csv')
Saccades_raw <- read_csv('data/saccades2018_2019.csv')

# variable types
Fixations_EL <- Fixations_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

Saccades_EL <- Saccades_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Direction = as.factor(Direction),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),
    Item = as.factor(Item)
  )

rm(Fixations_raw, Saccades_raw)
```



### Fixations & Saccades from Samples2 (own parameters)
I (Victor) have added a couple to the fixation to make it smooth with the logfiles. 
```{r}

# FIXATIONS
Fixations <- Samples2 %>% 
  group_by(ParticipantID, Trial, FixationNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionX = median(PositionX),
    PositionY = median(PositionY),
    Year = Year[1],
    PupilSize = median(PupilSize) #added. 
  ) %>% 
  filter(!is.na(FixationNum))

# SACCADES
Saccades <- Samples2 %>% 
  group_by(ParticipantID, Trial, SaccadeNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    SearchOrder = SearchOrder[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionXStart = min(PositionX),
    PositionXEnd = max(PositionX),
    AmplitudeX = max(PositionX)-min(PositionX),
    MeanVelocityX = median(MeanVelocityX),
    MeanAccellerationX = median(MeanAccellerationX),
    PositionYStart = min(PositionY),
    PositionYEnd = max(PositionY),
    AmplitudeY = max(PositionY)-min(PositionY),
    MeanVelocityY = median(MeanVelocityY),
    MeanAccellerationY = median(MeanAccellerationY),
    Year = Year[1] #added year
  ) %>% 
  filter(!is.na(SaccadeNum))
```

Merging logfiles and Fixations / Saccades
Now we are looking at a subset of the data (2018).
Do we agree that Riccardo/Fabio didn't find logfiles for 2019?

```{r}

# FIXATIONS 
# Fixations "Trial" column --> numeric for compatibility. 
Fixations$Trial <- as.numeric(Fixations$Trial)

#loading logfiles
logfiles <- read.csv("data/logfilesSE.csv")

#deleting 2018
Fixations$ParticipantID <- substring(Fixations$ParticipantID, 6, 20)

# SOCIAL ENGAGMENT
Fixations_SE_raw <- logfiles %>%
  full_join(Fixations, by = c("Task", "Trial", "ParticipantID")) %>%
  filter(Year == "2018") 

Fixations_SE <- Fixations_SE_raw %>%
  mutate(stim_person = ifelse(str_detect(Video, "m_"), "Male", "Female"),
         direct = ifelse(str_detect(Video, "dir"), "Direct", "Non-direct"),
         ost = ifelse(str_detect(Video, "_\\+o"), "Ostentive", "Non-ostentive")) 

#SACCADES 
#deleting 2018
Saccades$ParticipantID <- substring(Saccades$ParticipantID, 6, 20)

#Trial to numeric
Saccades$Trial <- as.numeric(Saccades$Trial)

# SOCIAL ENGAGMENT
Saccades_SE_raw <- logfiles %>%
  full_join(Saccades, by = c("Task", "Trial", "ParticipantID")) %>%
  filter(Year == "2018")
  #deleted the 2018 thing because that is 

Saccades_SE <- Saccades_SE_raw %>%
  mutate(stim_person = ifelse(str_detect(Video, "m_"), "Male", "Female"),
         direct = ifelse(str_detect(Video, "dir"), "Direct", "Non-direct"),
         ost = ifelse(str_detect(Video, "_\\+o"), "Ostentive", "Non-ostentive")) 

```

Preprocessing (Fixations_SE, Saccades_SE). 

```{r}
#Plot showing that something is weird. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#Which person is weird? 
which(Fixations_SE$StartTime > 7) #participantID 2018-1_2_f

#Max StartTime excluding the outlier 
Fixations_SE %>% 
  filter(!ParticipantID == "2018-1_2_f") %>%
  summarize(max(StartTime))

#Excluding every datapoint which has StartTime > 6.325
Fixations_SE <- Fixations_SE %>%
  filter(!StartTime > 6.325)

#Checking 
max(Fixations_SE$StartTime) #now excluded

#Plot showing that something is weird. 
ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#New column (condition)
Fixations_SE <- Fixations_SE %>%
  mutate(condition = paste0(direct, ost)) 

#Recoding to numbers
Fixations_SE$condition <- as.factor(as.numeric(as.factor(Fixations_SE$condition)))

#plot
ggplot(Fixations_SE, aes(condition, PupilSize)) +
  geom_violin() +
  geom_boxplot(width = .1)
  #direct & non-ostensive = 1 
  #direct & ostensive = 2
  #non-direct & non-ostensive = 3
  #non-direct & ostensive = 4

#making arousal variable
Fixations_SE <- Fixations_SE %>%
  mutate(arousal = ifelse(ParticipantGender == stim_person, "no", "yes"))

#### FIXES IN SACCADE ####

#Making a mean velocity column based on x & y velocity. 
Saccades_SE <- Saccades_SE %>%
  mutate(MeanVelocity = sqrt(MeanVelocityX^2 + MeanVelocityY^2))

#making arousal variable
Saccades_SE <- Saccades_SE %>%
  mutate(arousal = ifelse(ParticipantGender == stim_person, "no", "yes"))

#New column (condition)
Saccades_SE <- Saccades_SE %>%
  mutate(condition = paste0(direct, ost)) 

#Recoding to numbers
Saccades_SE$condition <- as.factor(as.numeric(as.factor(Saccades_SE$condition)))

#Same participant has longer trial (of course)
ggplot(Saccades_SE, aes(x = StartTime, y = MeanVelocity, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#Max StartTime excluding the outlier 
Saccades_SE %>% 
  filter(!ParticipantID == "2018-1_2_f") %>%
  summarize(max(StartTime))

#Excluding every datapoint which has StartTime > 6.426
Saccades_SE <- Saccades_SE %>%
  filter(!StartTime > 6.426)

#Checking 
max(Saccades_SE$StartTime) #now excluded

#Plot showing that something is weird. 
ggplot(Saccades_SE, aes(x = StartTime, y = MeanVelocity, color = ost)) +
  geom_smooth() +
  facet_grid(~direct)

#New column (condition)
Saccades_SE <- Saccades_SE %>%
  mutate(condition = paste0(direct, ost)) 

#Recoding to numbers
Saccades_SE$condition <- as.factor(as.numeric(as.factor(Saccades_SE$condition)))

#plot
ggplot(Saccades_SE, aes(condition, MeanVelocity)) +
  geom_boxplot()
  #direct & non-ostensive = 1 
  #direct & ostensive = 2
  #non-direct & non-ostensive = 3
  #non-direct & ostensive = 4
```

plotting distributions of dependent variables motivating models. 
p1: fix. duration log-normally distributed by participants (& across)
p2: Pupil size relatively normally dist. by part. (with a lot of noise)
p3: Pupil size bimodally distributed across participants (maybe issue)

```{r}



## same obtained across participants. 
p1 <- ggplot(Fixations_SE, aes(PupilSize, fill = "red")) + geom_density(alpha = 0.3) 
p1

## Check distributions of pupil size (per participant)
p2 <- ggplot(Fixations_SE, aes(PupilSize, fill = ParticipantID)) + geom_density(alpha = 0.3) #relatively normal (with noise). 
p2



## Check distribution of fixations duration (per participant)
p3 <- ggplot(Fixations_SE, aes(Duration, fill = "red")) + geom_density(alpha = 0.3) #super long tails - power-law. 
p3
#bi-modal 

## Check distribution of saccade velocity, power-law (also log-link)
p4 <- ggplot(Saccades_SE, aes(MeanVelocity, fill = "red")) + geom_density(alpha = 0.3)
p4

## combining pre-plots (include this.)
library(cowplot)
plot_grid(p1, p2, p3, p4, labels = c("bimodal", "normal (ish)", "power-law", "power-law")) 

```

# H1 MODEL (PUPIL SIZE)

```{r}

#Maximal interaction model with everything (non log-link)
h1_model1 <- lmer(PupilSize ~ ost * direct + StartTime + arousal + #include this
            (1 + arousal | ParticipantID) + #gender, familiarity individual diff
            (1 + ost * direct | ParticipantID) + #ost & dir as well. 
            (1 + StartTime | ParticipantID), 
            data = Fixations_SE,
            control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(h1_model1) #StartTime. 
anova(h1_model1)
tab_model(h1_model1, show.re.var = FALSE, show.icc = FALSE, pred.labels = TRUE)

qqnorm(residuals(h1_model1)) #looks terrible, one big outlier. 

```

H1 PLOTS (Pupil Size). 
p5: by condition & participantID. 
p6: by participantID & arousal. 
p7: by time & with directedness/ostension as fill/color. 
p7.1: by time & 

```{r}
#ridges showing why condition (ostension and directedness) is not significant. 
p5 <- ggplot(Fixations_SE, aes(x = PupilSize, y = ParticipantID)) +
  geom_density_ridges(alpha = 0.3, scale = 1.3, aes(fill = condition))  
p5 #more variance between participants than bewteen conditions. 

#ridges showing why gender/experimenter is not significant. 
p6 <- ggplot(Fixations_SE, aes(x = PupilSize, y = ParticipantID)) +
  geom_density_ridges(alpha = 0.3, scale = 1.3, aes(fill = arousal))  
p6

#ostension and directedness over time. 
p7 <- ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth(method = "lm", aes(fill = direct)) 
p7 #suggesting some effect. 

#loess, showing that the slopes are not in fact linear (only in the small world of the model). 
p7.1 <- ggplot(Fixations_SE, aes(x = StartTime, y = PupilSize, color = ost)) +
  geom_smooth(method = "loess", aes(fill = direct)) 
p7.1

#combining: INCLUDE THIS!
plot_grid(p5, p6, p7, p7.1, labels = c("condition", "arousal", "linear", "loess"))
```

H2: Fixation duration.
One big model with everything in it. 

```{r}

h2_model1 <- glmer(Duration ~ ost * direct + StartTime + arousal +
            (1 + arousal | ParticipantID) + #participants over stim. 
            (1 + ost * direct | ParticipantID) + #participants over ost & dir. 
            (1 + StartTime | ParticipantID), #participants over startTime. 
            family = gaussian(link=log), #log link because of distribution. 
            data = Fixations_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

#summary & qqnorm
summary(h2_model1) #male participants have larger pupil size.
qqnorm(residuals(h2_model1)) #looks terrible even with link = log. 
exp(-2.472130) #0.07811212 female intercept
exp(-2.472130 + 0.031310) #0.2280787 male intercept 
time_diff <- 0.08708941 - 0.08440488
time_diff
#looking better
tab_model(h2_model1, show.re.var = FALSE, show.icc = FALSE, pred.labels = TRUE)

#obtaining r2
library(MuMIn)
r.squaredGLMM(h2_model1)

#obtaining f-statistic
library("pbkrtest")

KRmodcomp(h2_model1)


```

H2 Plots:
p7: male participants fixate longer than female participants. 
p8: no big difference in fixation between conditions. 

```{r}

#duration by start time (color = participant ID)
p8 <- ggplot(Fixations_SE, aes(x = StartTime, y = Duration, color = condition)) +
  geom_smooth(method = "lm") 
p8 

#density plot of fixation duration by condition. 
p9 <- ggplot(Fixations_SE, aes(Duration, fill = condition)) + geom_density(alpha = 0.2)
p9 #no big difference between conditions. 

#combined, INCLUDE THIS
plot_grid(p8, p9, labels = c("duration by time", "Duration distribution"))
```

H3 saccade velocity

```{r}

h3_model1 <- glmer(MeanVelocity ~ ost * direct + StartTime + arousal +
            (1 + stim_person | ParticipantID) + #participants over stim. 
            (1 + ost * direct | ParticipantID) + #participants over ost & dir. 
            (1 + StartTime | ParticipantID), #participants over startTime. 
            family = gaussian(link=log), #log link because of distribution. 
            data = Saccades_SE,
            control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

summary(h3_model1)

```


