---
title: "P1_Visual Search"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(tidyverse, lme4, ggunchained, wesanderson)
```

Welcome,
this markdown is divided into following sections:
-preprocessing
-models
-results


# PREPROCESSING
Based on code from Fabio's day 4 Rmd

Outline:
a) working on raw data
-Loading raw data from the scanner
-Clean data (standardize dominant eye)
-Plotting raw data, 2 plots
-Event estimation
-Plotting raw data, while marking estimated events, 4 plots

b) working on event data
-appending Item
-appending Task

## A)
### loading data in right format
```{r}
# NO NEED TO RUN THIS! 
# DATASET FOR MODELING IS IMPORTED LATER ON. 
# FAST FORWARD TO NORMALIZATION

# loading data
Samples_raw <- read_csv('data/samples2018_2019.csv') %>%
  filter(Task == "VisualSearch")

# variable types
Samples_raw <- Samples_raw %>%
  mutate(
    ParticipantID = as.factor(ParticipantID),
    Trial = as.factor(Trial),
    DominantEye = ifelse(is.na(RightEye_PositionX), 'Left', 'Right'),
    RightEye_PositionX = as.numeric(RightEye_PositionX),
    RightEye_PositionY = as.numeric(RightEye_PositionY),
    RightEye_PupilSize = as.numeric(RightEye_PupilSize),
    RightEye_MeanVelocityX = as.numeric(RightEye_MeanVelocityX),
    RightEye_MeanVelocityY = as.numeric(RightEye_MeanVelocityY),
    RightEye_MeanAccellerationX = as.numeric(RightEye_MeanAccellerationX),
    RightEye_MeanAccellerationY = as.numeric(RightEye_MeanAccellerationY),
    LeftEye_PositionX = as.numeric(LeftEye_PositionX),
    LeftEye_PositionY = as.numeric(LeftEye_PositionY),
    LeftEye_PupilSize = as.numeric(LeftEye_PupilSize),
    LeftEye_MeanVelocityX = as.numeric(LeftEye_MeanVelocityX),
    LeftEye_MeanVelocityY = as.numeric(LeftEye_MeanVelocityY),
    LeftEye_MeanAccellerationX = as.numeric(LeftEye_MeanAccellerationX),
    LeftEye_MeanAccellerationY = as.numeric(LeftEye_MeanAccellerationY),
    Year = as.factor(Year),
    Task = as.factor(Task),
    ParticipantGender = as.factor(ParticipantGender),     
    SearchOrder = as.factor(SearchOrder),
    # sampling rate "is"" 1000 Hz, we're converting to miliseconds
    TrialTime = as.numeric(TrialTime*0.001),
    TotTrialLengthInSeconds = as.numeric(TotTrialLengthInSeconds)
  )
```


### cleaning of Samples
```{r}
Samples <- Samples_raw %>%
  # convert right- and left- data into a single column
  # note: position is in pixels, [0, 0] would be the bottom left corner
  # some machines do it differently
  mutate(
    PositionX = ifelse(DominantEye == 'Right', RightEye_PositionX, LeftEye_PositionX),
    PositionY = ifelse(DominantEye == 'Right', RightEye_PositionY, LeftEye_PositionY),
    PupilSize = ifelse(DominantEye == 'Right', RightEye_PupilSize, LeftEye_PupilSize),
    MeanVelocityX = ifelse(DominantEye == 'Right', RightEye_MeanVelocityX, LeftEye_MeanVelocityX),
    MeanVelocityY = ifelse(DominantEye == 'Right', RightEye_MeanVelocityY, LeftEye_MeanVelocityY),
    MeanAccellerationX = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationX, LeftEye_MeanAccellerationX),
    MeanAccellerationY = ifelse(DominantEye == 'Right', RightEye_MeanAccellerationY, LeftEye_MeanAccellerationY)
  ) %>% 
  # clear out the mess
  select(-c(RightEye_PositionX, RightEye_PositionY, RightEye_PupilSize, RightEye_MeanVelocityX, RightEye_MeanVelocityY, RightEye_MeanAccellerationX, RightEye_MeanAccellerationY, LeftEye_PositionX, LeftEye_PositionY, LeftEye_PupilSize, LeftEye_MeanVelocityX, LeftEye_MeanVelocityY, LeftEye_MeanAccellerationX, LeftEye_MeanAccellerationY))

rm(Samples_raw)
```


### plotting raw data
```{r}
x = subset(Samples, ParticipantID == '2019-2019-1_1_m11')
ggplot(x, aes(x = PositionX, y = PositionY, color = TrialTime)) +
  geom_point(size = 0.3) +
  ggtitle(paste('Sample data, subject:', x$ParticipantID, ', task:', x$Task )) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point() +
  scale_color_gradientn(colours = wes_palette("Zissou1", 21, type = "continuous"))
```


### estimation of events
```{r}
Samples2 <- Samples %>% 
  group_by(ParticipantID, Trial) %>%
  ## if velocity is < 20 degrees/sec, the eye is "not moving"
  mutate(
    # velocity goes from minus to plus depending on the direction, ORIGINAL = -20, 20
    LowVelocity = ifelse(
      as.numeric(MeanVelocityX)>-10 &
      as.numeric(MeanVelocityX)<10, 1, 0)
  ) %>%
  # assume movement during the first 9 rows (purely to make the code work)
  mutate(LowVelocity = coalesce(LowVelocity, 0)) %>%
  
  ## create rolling window for averaging, ORIGINAL = 20
  mutate(Event = (cumsum(LowVelocity) - lag(cumsum(LowVelocity), n = 5))/5) %>%
  
  ## the next lines are very hacky
  ## if average is > .9, count as fixation
  ## ORIGINAL = 0.75
  mutate(FixationNum = ifelse(Event >= 0.75, 1, 0),
         .count = ifelse(c(0, diff(FixationNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         FixationNum = ifelse(FixationNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  
  ## otherwise count as saccade
  mutate(SaccadeNum = ifelse(Event < 0.75, 1, 0),
         .count = ifelse(c(0, diff(SaccadeNum)) == 1, 1, 0),
         .count = coalesce(.count, 0),
         .groups = cumsum(.count),
         SaccadeNum = ifelse(SaccadeNum == 1, .groups, NA)) %>%
  select(-c(.count, .groups)) %>% 
  mutate(Event = ifelse(Event >= 0.75, 'Fixation', 'Saccade'))

rm(Samples)
```


### plotting event estimation: spatial
```{r}
# PLOT FOR TRIAL 1
ggplot(subset(Samples2, ParticipantID == '2019-2019-1_1_m11' & Trial == 1 & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Sample data with event estimation, trial', x$Trial)) +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  theme_janco_point()
  

# FACET PLOT FOR EVERY TRIAL
ggplot(subset(Samples2, ParticipantID == '2019-2019-1_1_m11' & !is.na(Event)),
       aes(x = PositionX, y = PositionY, color = Event)) +
  geom_point(size = 0.3) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Sample data with event estimation, all trials') +
  xlab(paste('Position X (', x$DominantEye, ')', sep = '')) +
  ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  xlim(0,1500) + ylim(0,1250) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


### plotting event estimation: temporal & spatial
```{r}
ggplot(subset(Samples2, ParticipantID == '2019-2019-1_1_m11' & Trial == 1 & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.6) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle(paste('Time sequence plot, trial', x$Trial)) +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  theme_janco_point()

ggplot(subset(Samples2, ParticipantID == '2019-2019-1_1_m11' & !is.na(Event)),
       aes(x = TrialTime, y = PositionY, color = Event)) +
  geom_point(size = 0.5) +
  ## the following lines are for aesthetic purposes only 
  scale_color_manual(values = c('#6291D6', '#F0B952'), name = 'Event') +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ggtitle('Time sequence plot, all trials') +
  xlab('Time (s)') + ylab(paste('Position Y (', x$DominantEye, ')', sep = '')) +
  facet_wrap(~Trial, ncol = 4) +
  theme_janco_point()
```


## B
### Fixations & Saccades from Samples2 (own parameters)
```{r}
# FIXATIONS
Fixations <- Samples2 %>% 
  group_by(ParticipantID, Trial, FixationNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    SearchOrder = SearchOrder[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionX = median(PositionX),
    PositionY = median(PositionY)
  ) %>% 
  filter(!is.na(FixationNum))

# SACCADES
Saccades <- Samples2 %>% 
  group_by(ParticipantID, Trial, SaccadeNum) %>% 
  summarize(
    Task = Task[1],
    ParticipantGender = ParticipantGender[1],
    SearchOrder = SearchOrder[1],
    StartTime = min(TrialTime),
    EndTime = max(TrialTime),
    Duration = max(TrialTime)-min(TrialTime),
    PositionXStart = min(PositionX),
    PositionXEnd = max(PositionX),
    AmplitudeX = max(PositionX)-min(PositionX),
    MeanVelocityX = median(MeanVelocityX),
    MeanAccellerationX = median(MeanAccellerationX),
    PositionYStart = min(PositionY),
    PositionYEnd = max(PositionY),
    AmplitudeY = max(PositionY)-min(PositionY),
    MeanVelocityY = median(MeanVelocityY),
    MeanAccellerationY = median(MeanAccellerationY)
  ) %>% 
  filter(!is.na(SaccadeNum))
```


### append itmes in visual searching
```{r}
ConditionsVisualSearch <- read_delim('data/ForagingOrder.csv', ';') %>%
  mutate(Item = ifelse(Item == "doplhins.jpg", "dolphins.jpg", as.character(Item)),
         Trial = as.factor(Trial),
         SearchOrder = as.factor(SearchOrder))

Fixations_mrg <- full_join(Fixations, ConditionsVisualSearch, 
                   by = c('Trial', 'SearchOrder'), all = TRUE)

Saccades_mrg <- full_join(Saccades, ConditionsVisualSearch, 
                   by = c('Trial', 'SearchOrder'), all = TRUE)

Fixations_VS <- Fixations_mrg %>%
  filter(Task == "VisualSearch")

Saccades_VS <- Saccades_mrg %>%
  filter(Task == "VisualSearch")

rm(ConditionsVisualSearch, Fixations, Saccades, Fixations_mrg, Saccades_mrg)
```

### append condition into VS
```{r}
### getting the conditions in there
### number of combinations = 12 participants * 2 conditions * 5 trials - 6 missing trials of the first participant
condition_task <- read_csv("data/fixations2018_2019.csv") %>%
  filter(Task != "SocialEngagement") %>%
  mutate(Trial = as.character(Trial),
         Task = as.character(Task),
         ParticipantID = as.character(ParticipantID)) %>%
  select(Task, Trial, ParticipantID) %>%
  unique.data.frame()

Fixations_VS_2 <- Fixations_VS %>%
  ungroup() %>%
  mutate(Task = as.character(Task),
         Trial = as.character(Trial),
         ParticipantID = as.character(ParticipantID)) %>%
  mutate(ParticipantID = substring(ParticipantID, 6, 17)) %>%
  select(-Task) %>%
  full_join(condition_task)

Saccades_VS_2 <- Saccades_VS %>%
  ungroup() %>%
  mutate(Task = as.character(Task),
         Trial = as.character(Trial),
         ParticipantID = as.character(ParticipantID)) %>%
  mutate(ParticipantID = substring(ParticipantID, 6, 17)) %>%
  select(-Task) %>%
  full_join(condition_task)
```

### clean global environment
```{r}
rm(Saccades_VS, Fixations_VS, condition_task)

Fixations_VS_2 <- Fixations_VS_2 %>%
  mutate(ParticipantID = as.factor(ParticipantID),
         Trial = as.factor(Trial),
         FixationNum = as.factor(FixationNum),
         Task = as.factor(Task))

Saccades_VS_2 <- Saccades_VS_2 %>%
  mutate(ParticipantID = as.factor(ParticipantID),
         Trial = as.factor(Trial),
         SaccadeNum = as.factor(SaccadeNum),
         Task = as.factor(Task))
```


######
ready made data
######
```{r}
Fixations_VS_2 <- read_csv("data/fixations_VS_2.csv") %>%
  mutate_if(is.character, as.factor) %>%
  mutate(SearchOrder = factor(SearchOrder),
         Trial = factor(Trial))

Saccades_VS_2 <- read_csv("data/saccades_VS_2.csv") %>%
  mutate_if(is.character, as.factor) %>%
  mutate(SearchOrder = factor(SearchOrder),
         Trial = factor(Trial))
```



## NORMALISATION
```{r}
# x / max(x)ing
Fixations_VS_3 <- Fixations_VS_2 %>%
  mutate(Duration = Duration / max(Duration),
         StartTime = StartTime / max(StartTime))

Saccades_VS_3 <- Saccades_VS_2 %>%
  mutate(Duration = Duration / max(Duration),
         StartTime = StartTime / max(StartTime),
         # amplitudes on different axes
         AmplitudeX = AmplitudeX / max(AmplitudeX),
         AmplitudeY = AmplitudeY / max(AmplitudeY),
         # combined amplitude
         Amplitude_comb = sqrt(AmplitudeX^2 + AmplitudeY^2),
         Amplitude_comb = Amplitude_comb / max(Amplitude_comb),
         # index
         SaccadeNum = as.numeric(as.character(SaccadeNum)),
         SaccadeNum = SaccadeNum / max(SaccadeNum))

```


# FINDING STAR TRIALS
## generate scanpaths for everyone
```{r}
plot_n_save2 <- function(participant) {
    one_person <- Fixations_VS_3 %>%
      filter(ParticipantID == as.character(participant)) %>%
      filter(Task == "VisualSearch_Star")
    
    plot <- ggplot(one_person, aes(x = PositionX, 
                                   y = 1081 - PositionY)) +
      geom_path(alpha = .6, color = '#F0B952') +
      geom_point(size = one_person$Duration*5, 
                 alpha = .6, color = '#6291D6', stroke = 1, shape = 21) +
      # repelling text in plots: fuck that's good
      # ggrepel::geom_text_repel(aes(label = FixationNum), size = 3, color = 'black') +
      xlim(0,1920) + ylim(0,1080) +
      labs(title = paste0("Visual search - ",
                          as.character(unique(one_person$ParticipantID))),
           subtitle = paste0("Seach order: ", 
                             as.character(unique(one_person$SearchOrder)))) +
      theme_bw() + theme(legend.position = 'none') +
      facet_wrap(~Item)
    
    ggsave(sprintf("pics/star_trials/file%s.png", 
                   unique(one_person$ParticipantID)), plot)
}

participants <- as.data.frame(table(Fixations_VS_3$ParticipantID)) %>%
  filter(Freq != 0)

participants <- as.character(participants$Var1)

# test
# plot_n_save2(participants[1])

# loop - WATCH OUT, PRODUCES A LOT OF FILES
lapply(participants, plot_n_save2)
```



# MODELS

```{r}
Model1 <- glmer(Duration ~ Task + StartTime + 
            (1 + Task | Item) + 
            (1 + Task + StartTime | ParticipantID),
            family = gaussian(link="log"), data = Fixations)

Model2 <- glmer(NSaccades ~ Task + StartTime + 
            (1 + Task | Item) + 
            (1 + Task + StartTime | ParticipantID),
            family = gaussian(link="log"), data = Saccades)

Model3<- glmer(Amplitude ~ Task + StartTime + 
            (1 + Task | Item) + 
            (1 + Task + StartTime | ParticipantID),
            family = gaussian(link="log"), data = Saccades)
```


# RESULTS

```{r}

```

